---
title: "**Factors Related to Parental Involvement in School**"
subtitle: "Data Science Institute, American University"
author: "Tanesia Beverly, Andy Zukerberg, Erin Murrock"
date: "1/11/2020"
output: 
  powerpoint_presentation
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning=FALSE, message=FALSE)
setwd("C:/Users/trbeverly/Documents/")
load("parental_involvement.Rdata")

```

## **Parental Involvement In School** 

- Parent involvement is related to success of children in school
- “The research overwhelmingly demonstrates that parent involvement in children's learning is positively related to achievement. Further, the research shows that the more intensively parents are involved in their children's learning, the more beneficial are the achievement effects. This holds true for all types of parent involvement in children's learning and for all types and ages of students.” (cotton and Reed, 1989)   


## **Study Purpose** 

- Bullet 1
- Bullet 2
- Bullet 3

## **PFI Questionnaire**

- Parent and Family Involvement in Education
- Conducted in 2016 by the National Center for Education Statistics
- Collected every two or three years
- 14,075 completed interviews
- Parents respond about a selected child
- Includes homeschooled and enrolled children

## **Research Questions and Hypotheses** 

- What are predictors of parent involvement in school?
- Bullet 2
- Bullet 3

## **Data Cleaning**

- Recoded data to remove valid skips 
- Converted some variables to 0 for no and 1 for yes to facilitate counting
- Created new variables for family activities, school involement and satisfaction with school 
- Removed homeschooled children


## **Participants/Demographics**

- Bullet 1
- Bullet 2
- Bullet 3

## **Variables of Interest**
Age2015= Age of child as of December 31st, 2015

Hhtotalxx= number of people in the household

Pargradex= parent or guardians highest level of education

Ttlhhinc= total household income, categorical

Scpubpri= child is enrolled in a public or private school

Sneighbrx= parent moved to neighborhood for school

Sefuturex=parent’s expectation for child’s future education (not finish HS to graduate school)

RACEETH2- recoded race and ethnicity of child

## **family activites measures**

FA-count of at home activities. Comprised of:
FOSTORY2X (told child a story past week), focrafts(did arts and crafts past week), fogames (played games with child past week), fobuildx (worked on prokect with child past week), fosport(played sports with child past week), forespon (discussed time management with child past week), fohistx (talked about family history with child past week), folibrayx (took child to library past month, fobookstx (took child to bookstore past month), foconcrtx (took child to concert past month), fomuseumx (took child to museum past month), fozoox (took child to zoo past month), fogroupx (attended event with child past month), fosprtevx (attend athletic event with child past month)

## slide for school satisfaction survey
SSatis- satisfaction with school. Comprised of:
FCSHOOL (satisfaction with school child attends, FCTEACHR (satisfaction with teachers), FCSTDS (satisfaction with academic standards), FCORDER (satisfaction with order and discipline at school), FCSUPPRT (satisfaction with way school interacts with parents

## **Analysis**

- Descriptive Statistics
- Dimension Reduction (EFA)
- Regression Analysis

## **Results: Dimension Reduction**

- Bullet 1
- Bullet 2
- Bullet 3

## **Variables for regression, means, etc Descriptives**

- Bullet 1
- Bullet 2
- Bullet 3

## **Outcome Variable**
FS- count of school activities parent participated in. 
Comprised of:	
FSSPORTX (attended school or class event, science fair, etc.), FSVOL(volunteered at school), FSMTNG(attended a general school meeting), FSPTMTNG (attended a PTA meeting, FSATCNFN ( gone to a parent teacher conference), FSFUNDRS (participated in fundraising for the school), FSCOMMTE (served on a school committee), FSCOUNSLR (met with the guidance counselor in person).


## **Multiple Regression Model**
Residuals:
    Min      1Q  Median      3Q     Max 
-5.7709 -1.0177  0.0583  1.0501  5.4114 

Coefficients:
             Estimate Std. Error t value Pr(>|t|)    
(Intercept)  3.224210   0.116115  27.767   <2e-16 ***
AGE2015     -0.070933   0.004016 -17.662   <2e-16 ***
hhtotalxx    0.011937   0.010832   1.102   0.2705    
pargradex    0.199829   0.014607  13.680   <2e-16 ***
ttlhhinc     0.087108   0.005825  14.955   <2e-16 ***
FA           0.260147   0.008965  29.018   <2e-16 ***
SSatis      -0.054204   0.004512 -12.013   <2e-16 ***
public      -0.768399   0.042936 -17.896   <2e-16 ***
sneighbrx    0.073768   0.034558   2.135   0.0328 *  
sefuturex    0.125544   0.011685  10.744   <2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
Residual standard error: 1.524 on 13513 degrees of freedom
Multiple R-squared:  0.2726,	Adjusted R-squared:  0.2721 
F-statistic: 562.7 on 9 and 13513 DF,  p-value: < 2.2e-16

## **REgression output
\tiny
```{r echo = FALSE}

fit <-lm(FS ~ AGE2015 + hhtotalxx + AGE2015 + pargradex + ttlhhinc + FA + SSatis + public + sneighbrx + sefuturex, data=parinv) 

summary(fit)

```

## **Poisson Model**
```{r echo=FALSE}
posfit <- glm(FS ~ AGE2015 + hhtotalxx + AGE2015 + pargradex + ttlhhinc + FA + SSatis + public + sneighbrx + sefuturex, data=parinv, family = poisson())
summary(posfit)
```

##**Poisson Regression**
Deviance Residuals: 
    Min       1Q   Median       3Q      Max  
-3.4397  -0.5443   0.0289   0.5053   2.6498  
Coefficients:
             Estimate Std. Error z value Pr(>|z|)    
(Intercept)  1.095827   0.038737  28.289   <2e-16 ***
AGE2015     -0.016818   0.001307 -12.866   <2e-16 ***
hhtotalxx    0.004260   0.003602   1.182    0.237    
pargradex    0.051818   0.004926  10.520   <2e-16 ***
ttlhhinc     0.022283   0.001959  11.372   <2e-16 ***
FA           0.066241   0.002983  22.208   <2e-16 ***
SSatis      -0.014556   0.001554  -9.367   <2e-16 ***
public      -0.162918   0.012828 -12.700   <2e-16 ***
sneighbrx    0.015885   0.011133   1.427    0.154    
sefuturex    0.035777   0.004083   8.762   <2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
(Dispersion parameter for poisson family taken to be 1)
    Null deviance: 13187  on 13522  degrees of freedom
Residual deviance: 10243  on 13513  degrees of freedom
AIC: 52140
Number of Fisher Scoring iterations: 4

## **Discussion**

- Bullet 1
- Bullet 2
- Bullet 3

## **Future Implications**

- Bullet 1
- Bullet 2
- Bullet 3

## **THANK YOU!**
